### 2018-02-05 paper/ICML/related_work.tex

\section{Related Work}

\begin{itemize}
    \item Hyperparameter optimization: random, bayesian opt, bandit methods
    \item distillation techniques
    \item post-training compression techniques
    \item group sparsity, non-parametric neural networks
    \item training dynamics paper: first overfitting and then randomization?
  \end{itemize}