### 2018-02-05 paper/ICML/discussion.tex

\section{Discussion}

\begin{itemize}
  \item Where does the method shine?
  \item Side-effects of method: smaller networks, time to train
  \item Potential extensions/limitations
\end{itemize}