### 2018-02-09 paper/ICML/related_work.tex
Resolve even more references

\section{Related Work}

\begin{itemize}
  \item Hyperparameter optimization: random, bayesian opt, bandit methods
  \item distillation techniques
  \item post-training compression techniques
  \item group sparsity, non-parametric neural networks
  \item training dynamics paper: first overfitting and then randomization?, \gl{Here is the ref, if you can introduce it in the flow \cite{Shwartz-Ziv2017}}
\end{itemize}

Given the importance of network structure, many techniques have been proposed
to find the best network structure for a given learning task.  These techniques
broadly fall into four categories: hyperparameter optimization strategies,
post-training model compression for inference as well as model simplification,
techniques to resize models during training, and automated architecture search
methods.

The most popular techniques for hyperparameter optimization include simple
methods such as random search~\cite{BergstraJAMESBERGSTRA2012} which have been
shown to work suprisingly well compared to more complex methods such as those
based on Bayesian optimization~\cite{Snoek12}.  Techniques such
as~\cite{Snoek12} model generalization performance via Gaussian
Processes~\cite{GaussianProcesses} and select hyperparameter combinations that
come from uncertain areas of the hyperparameter space.  Recently, methods based
on bandit algorithms (e.g. ~\cite{li2016hyperband, jamieson2016}) have also
become a popular way to tune hyperparameters.  As noted before, all of the
above techniques require many tens to hundreds of models to be trained, making
this process computationally inefficient and slow.

In contrast with hyperparameter tuning methods, some methods such as
DeepCompression~\cite{han2015deepcompression} seek to compress the network to
make inference more efficient.  This is accomplished by pruning connections and
quantizing weights.  On similar lines, multiple techniques such
as~\cite{romero2014fitnets, hinton2015distilling} have been proposed for
distilling a network into a simpler network or a different model.  Unlike our
technique which works during training, these techiques are used after training
and it would be interesting to apply them to ShrinkNets as well. \cite{Abadi2016b} share the common goal of removing entire blocks of parameter to maintain dense matrices, however their method only applies to convolutional layers.

The techniques closed to our work are those based on group sparsity such as
~\cite{Scardapane2017}, and those like ~\cite{Philipp} who grow and shrink
dynamically the size of the network during training.

Finally, there has also been recent work in automatically learning model
architecture through the use of genetic algorithms and reinforcement learning
techniques~\cite{DBLP:journals/corr/ZophL16, zoph2017learning}.  These
techniques are focused more on learning higher-level architectures (e.g.
building blocks for neural network architectures) as opposed to learning
network size.