### 2018-02-09 paper/ICML/theory.tex

%\subsection{Relation to Group Sparsity}
\noindent\textbf{Relation to Group Sparsity (LASSO): } ShrinkNets removes neurons,
i.e., inputs and outputs of layers. For a fully connected layer defined as:

\begin{equation} \label{fully_connected}
  f_{\bm{A}, \bm{b}}(\bm{x})=a(\bm{Ax + b})
\end{equation}

Removing an input neuron $j$ is equivalent to have $\left(\bm{A}^T\right)_j =
\bm{0}$. Removing an output neuron $i$ is the same as having $\bm{A}_i = \bm{0}$
and $\bm{b}_i = 0$. Solving optimization problems while trying to set entire
group of parameters to zero is the goal of group sparsity regularization
\cite{}. For any partitioning of the set of parameter defining a model in $p$
groups: $\bm{\theta} = \bigcup_{i=1}^P \bm{\theta}_i$ we define it the following
way:

\begin{equation}
  \Omega_\lambda^{gp} = \lambda \sum_{i=1}^p \sqrt{\#\bm{\theta_i}} \norm{\bm{\theta_i}}_2 \\
\end{equation}

In fully-connected layers, the groups are either: columns of
$\bm{A}$ if we want to remove inputs, or rows of $\bm{A}$ and the corresponding
entry in $\bm{b}$ if we want to remove outputs. For simplicity, we focus
our analysis in the simple one-layer case. In this case, filtering outputs does
not make sense, this is why we will only consider the former case. The
group sparsity regularization then becomes:

\begin{equation} \label{group_sparsity_regularization}
  \Omega_\lambda^{gp} = \lambda \sum_{j=1}^p \norm{\bm{\left(A^T\right)_j}}_2 \\
\end{equation}

Because $\forall i, \#\bm{\theta}_i = n$, \ra{is \# the general way of expressing
cardinality? why not |x|?}to make the notation simpler, we
embedded $\sqrt{n}$ inside $\lambda$.

Group sparsity and ShrinkNets try to achieve the same goal. We discuss next how
they are related to each other. First let's recall the two problems. The
original ShrinkNet problem is:

\begin{equation}
  \min_{\bm{A}, \bm{\beta}} \norm{\bm{y} - \bm{A}\diag{\bm{\beta}}\bm{x}}_2^2 + \lambda \norm{\bm{\beta}}_1
\end{equation}

And the Group Sparsity problem is:

\begin{equation}
  \min_{\bm{A}} \norm{\bm{y} - \bm{A}\bm{x}}_2^2 + \Omega_\lambda^{gp}
\end{equation}

We can prove that under the condition: $\forall j\in \intint{1, p},
\norm{\left(\bm{A}^T\right)_j}_2 = 1$ the two problems are equivalent
(\cref{gps_equivalence}). However if we relax this constraint then shrinknet
becomes non-convex and has no global minimum
(\cref{unconstrained_non_convex,unconstrained_shrinknet_no_min}). Fortunately,
by adding an extra term to the ShrinkNet regularization term we can proove that:

\begin{equation}
  \min_{\bm{A}, \bm{\beta}} \norm{\bm{y} - \bm{A}\diag{\bm{\beta}}\bm{x}}_2^2 + \Omega_\lambda^s + \lambda_2\norm{A}_p^p
\end{equation}

has global minimums for all $p>0$ (\cref{shrinknet_regularized_minimum}).
This is the reason we defined the \textit{regularized ShrinkNet penalty} earlier
this way:\ra{the notation is slightly different, better to make it consistent}

\begin{equation}
  \Omega_{\lambda, \lambda_2, p}^{rs} = \lambda\norm{\bm{\beta}}_1 + \lambda_2\norm{\bm\theta}_p^p
\end{equation}

In practice we observed that $p=2$ or $p=1$ are good choice, while the latter
will also introduce additional sparsity in the parameters.