### 2018-02-09 paper/ICML/related_work.tex
related work, makefile

%!TEX root=paper.tex
\section{Related Work}

\begin{itemize}
  \item post-training compression techniques -- brain damage , 
  \item group sparsity e.g., \cite{Scardapane2017} and non-parametric neural networks -- 
  \item training dynamics paper: first overfitting and then randomization?, \gl{Here is the ref, if you can introduce it in the flow \cite{Shwartz-Ziv2017}}
\end{itemize}

Given the importance of network structure, many researchers have explored the
problem of finding the best network structure for a given learning task.  
The proposed techniques broadly fall into five categories: 
random search andbrute force search,
hyperparameter optimization,
model compression after training,
resizing models during training, and 
automated architecture search methods.

The most popular techniques for hyperparameter optimization include simple
methods such as random search~\cite{BergstraJAMESBERGSTRA2012} which have been
shown to work suprisingly well compared to more complex methods such as those
based on Bayesian optimization~\cite{Snoek12}.
Brute force search of network sizes is also become more practical due to faster
and more powerful hardware~\cite{molchanov2016pruning}.
The more complex methods for hyperparameter optimization include techniques 
such as~\cite{Snoek12} which model generalization performance via Gaussian
Processes~\cite{GaussianProcesses} and select hyperparameter combinations that
come from uncertain areas of the hyperparameter space.  
Recently, methods based
on bandit algorithms (e.g. ~\cite{li2016hyperband, jamieson2016}) have also
become a popular way to tune hyperparameters.
As noted before, all of the
above techniques require many tens to hundreds of models to be trained, making
this process computationally inefficient and slow.
More practically, the hyperparameter optimization literature does not evaluate
their methods on network size and instead focuses on optimization hyperparameters
such as learning rates and weight decay parameters.

In contrast with size search techniques, there are techniques that 
reduce network size {\it once the network has been traiend}.
Popular techniques in this class include optimal brain damage~\cite{Cun} and
compression techniques such as DeepCompression~\cite{han2015deepcompression}.
Optimal brain damage identifies connections in a network that are unimportant
and then prunes these connections.
DeepCompression takes this one step further and in addition to pruning connections,
it quantizes weights to make inference extremely efficient.
% In contrast with hyperparameter tuning methods, some methods such as
% DeepCompression~\cite{han2015deepcompression} seek to compress the network to
% make inference more efficient.  This is accomplished by pruning connections and
% quantizing weights.  
A different vein of work such as ~\cite{romero2014fitnets, hinton2015distilling}
proposes techniques for distilling a network into a simpler network or 
a different model.  Unlike our
technique which works during training, these techiques are used after training
and it would be interesting to apply them to ShrinkNets as well. 
\cite{Abadi2016b} share the common goal of removing entire blocks of parameter to maintain dense matrices, however their method only applies to convolutional layers.

The techniques closest to our proposed method are those based on group sparsity 
such as ~\cite{Scardapane2017}, and those like~\cite{Philipp} who grow and shrink
dynamically the size of the network during training.

\cite{Scardapane2017} presents a method that also deactivates neurons using a 
loss function based on group-sparsity.
However, the exact details of how their method works are not given, and 
their experimental results (on a small, fully connected network), 
are substantially worse than ours as shown in Section~\ref{sec:experiments}.
~\cite{Philipp} propose a method called Adaptive Radial-Angular Gradient Descent
that adds neurons on the fly and removes neurons via an $l_2$ penalty.
However, this approach requires a new optimizer and takes longer to converge 
compared to ShrinkNets.

Finally, there has also been recent work in automatically learning model
architecture through the use of genetic algorithms and reinforcement learning
techniques~\cite{DBLP:journals/corr/ZophL16, zoph2017learning}.  These
techniques are focused more on learning higher-level architectures (e.g.
building blocks for neural network architectures) as opposed to learning
network size.