### 2018-02-09 paper/ICML/related_work.tex
pass over related work

%!TEX root=paper.tex
\section{Related Work}

%\begin{itemize}
%  \item post-training compression techniques -- brain damage , 
%  \item group sparsity e.g., \cite{Scardapane2017} and non-parametric neural networks -- 
%  \item training dynamics paper: first overfitting and then randomization?, \gl{Here is the ref, if you can introduce it in the flow \cite{Shwartz-Ziv2017}}
%\end{itemize}

There are several lines of work related to finding a good network structure. We
describe them next:

%Given the importance of network structure, many researchers have explored the
%problem of finding the best network structure for a given learning task.  The
%proposed techniques broadly fall into five categories: random search andbrute
%force search, hyperparameter optimization, model compression after training,
%resizing models during training, and automated architecture search methods.

\noindent\textbf{Hyperparameter optimization techniques: }The most popular
techniques for hyperparameter optimization include simple methods such as random
search~\cite{BergstraJAMESBERGSTRA2012} which have been shown to work
suprisingly well compared to more complex methods such as those based on
Bayesian optimization~\cite{Snoek12}.  Brute force search of network sizes is
also become more practical due to faster and more powerful
hardware~\cite{molchanov2016pruning}.  The more complex methods for
hyperparameter optimization include techniques such as~\cite{Snoek12} which
model generalization performance via Gaussian Processes~\cite{GaussianProcesses}
and select hyperparameter combinations that come from uncertain areas of the
hyperparameter space. Recently, methods based on bandit algorithms (e.g.
~\cite{li2016hyperband, jamieson2016}) have also become a popular way to tune
hyperparameters. With respect to our work, these methods require many tens of
hundreds of models to be trained, which makes the process computationally
inefficient and slow. In contrast, with ShrinkNets all the hyperparameters
related to network size are reduced to a single parameter, as we explain next.
More importantly, none of the hyperparameter optimization methods focuses on
finding small networks, which is a crucial property of ShrinkNets, necessary to
achieve good inference times.

%As noted before, all of the above techniques require many tens
%to hundreds of models to be trained, making this process computationally
%inefficient and slow.  More practically, the hyperparameter optimization
%literature does not evaluate their methods on network size and instead focuses
%on optimization hyperparameters such as learning rates and weight decay
%parameters.

\noindent\textbf{Model Compression: }Model compression techniques focus on
reducing the model size \emph{after} training, in contrast to ShrinkNets, which
reduces it \emph{while} training. 
Optimal brain damage~\cite{Cun} identifies connections in a network that are
unimportant and then prunes these connections.
DeepCompression~\cite{han2015deepcompression} takes this one step further and in
addition to pruning connections, it quantizes weights to make inference
extremely efficient.  A different vein of work such as ~\cite{romero2014fitnets,
hinton2015distilling} proposes techniques for distilling a network into a
simpler network or a different model. Because these techniques work after
training, they are orthogonal and complementary to ShrinkNets. 
%Unlike our technique which works during
%training, these techiques are used after training and it would be interesting to
%apply them to ShrinkNets as well. 
%\cite{Abadi2016b} share the common goal of
%removing entire blocks of parameter to maintain dense matrices, however their
%method only applies to convolutional layers.

\noindent\textbf{Auto-ML: }There is work that focuses on automatically learning
model architecture through the use of genetic algorithms and reinforcement
learning techniques~\cite{DBLP:journals/corr/ZophL16, zoph2017learning}. These
techniques are focused on learning higher-level architectures (e.g.,
building blocks for neural network architectures). In particular, they do not
focus on finding small but well-performing networks for inference, which is the
goal of ShrinkNets.

\noindent\textbf{Dynamically Sizing Networks: }The techniques closest to our
proposed method are those based on group sparsity such as
~\cite{Scardapane2017}, and those like~\cite{Philipp} who grow and shrink
dynamically the size of the network during training.  \cite{Scardapane2017}
presents a method that also deactivates neurons using a loss function based on
group-sparsity.  However, the exact details of how their method works are not
given, and their experimental results (on a small, fully connected network), are
substantially worse than ours as shown in Section~\ref{sec:experiments}.
~\cite{Philipp} propose a method called Adaptive Radial-Angular Gradient Descent
that adds neurons on the fly and removes neurons via an $l_2$ penalty.  However,
this approach requires a new optimizer and takes longer to converge compared to
ShrinkNets.