### 2018-05-16 paper/NIPS/impl.tex

%!TEX root=paper.tex

\section{ShrinkNets in Practice}

In this section we discuss practical aspects of \shrink{}, including 
neuron removal  and several optimizations.

\paragraph{On-The-Fly Neuron Removal}
\label{neuron_killing}

Switch layers are initialized with weights sampled from $\mathcal{N}(0,1)$;
their values change as part of the training process so as to switch \emph{on}
or \emph{off} neurons. Using gradient descent, it is very unlikely that
components of $\bm{\beta}$ will eventually reach exactly $0$. In most cases,
irrelevants neurons will see their \swl oscillate between small positive and
negative values influenced solely by the $L_1$ penalty. Our goal is
to detect this situation and effectively set these components to $0$ exactly. We
evaluated multiple screening strategies but the most efficient and flexible one
was the \textbf{Sign variance strategy}: At each update we measure the sign of
each component of $\bm{\beta}$ ($-1$ or $1$). We maintain two metrics: the
exponential moving average (EMA) of its mean and variance. When the variance exceed a
predefined threshold we consider the value has been oscillating for a
sufficiently long and that the neuron does not participate to the output with a
high likelyhood, so we effectively deactivate it. This strategy is parametrized
by two hyper-parameters, the threshold but also the intertial of the EMA.

% When neurons are deactivated by switch layers, we need
% to remove them to actually shrink the network size.
% In the following we discuss three strategies to decide when to deactivate neurons. 
% %Since after removing a neuron it will no longer be able to contribute to  prediction accuracy, we must devise strategies that remove neurons while remaining robust to changing gradient values. We observe three strategies:
% 
% \paragraph{Threshold strategy} Under this strategy neurons are removed based on
% a fixed threshold expressed in terms of its absolute value. This is the method
% used by deep compression \cite{Han2015}. We found this method is very sensitive to the initialization 
% and not robust to noise in the gradients---which occurs commonly when training networks with
% stochastic gradient descent---because the threshold depends on the scale of
% each weight.
% 
% \textbf{Sign change strategy}: Under this strategy neurons are deactivated when the
% weight value changes its sign. As before this strategy is also sensitive to gradient noise; if we sample the gradient in a way that is not fully representative of the dataset we might experience one-time zero crossing which could wrongly deactivate a neuron.

% \textbf{Sign variance strategy}: Instead of removing on the first zero
% crossing, we can instead measure the exponential moving variance of the sign
% ($-1, 1$) of each parameter in the \textit{switch layer}. When the value of
% the exponential moving variances goes over a predefined threshold, we consider
% the neuron deactivated. This strategy introduces
% two extra parameters, one to control the behavior of the inertia of the
% variance, and the threshold. 
% However, we found that this strategy performs the best in practice and is used throughout our evaluation section. 
% 
%The last two strategies perform better than the threshold strategy.  Both effectively allow us to shrink the network by varying $\lambda$, and result in networks that are both small and perform well. We use the third one in our evaluation section.

\paragraph{Preparing for Inference} With \shrink we obtain
reduced-sized networks during training, which is the first steps towards faster
inference. This networks are readily available for inference. However, because
they include switch layers---and therefore more parameters---they introduce
unnecessary overhead at inference time. To avoid this overhead, we reduce the
network parameters by combining each switch layer with its respective network
layer by multiplying the respective parameters before emitting the final trained
network. As a result, the final network is a dense network without any switching layers.

\paragraph{Neural Garbage Collection} \shrink decides on-the-fly which
neurons to deactivate. Since \shrink deactivate a large fraction of neurons, we must
dynamically remove these neurons at runtime to not unnecessarily impact network
training time. We implemented a neural garbage collection method as part of our
library which takes care of updating the necessary network layers as well as
updating optimizer state to reflect the neuron removal.