### 2018-05-16 paper/NIPS/paper.tex

\documentclass{article}

% if you need to pass options to natbib, use, e.g.:
% \PassOptionsToPackage{numbers, compress}{natbib}
% before loading nips_2018

% ready for submission
\usepackage[preprint,nonatbib]{nips_2018}

% to compile a preprint version, e.g., for submission to arXiv, add
% add the [preprint] option:
% \usepackage[preprint]{nips_2018}

% to compile a camera-ready version, add the [final] option, e.g.:
% \usepackage[final]{nips_2018}

% to avoid loading the natbib package, add option nonatbib:
% \usepackage[nonatbib]{nips_2018}

\usepackage[utf8]{inputenc} % allow utf-8 input
\usepackage[T1]{fontenc}    % use 8-bit T1 fonts
\usepackage{hyperref}       % hyperlinks
\usepackage{url}            % simple URL typesetting
\usepackage{booktabs}       % professional-quality tables
\usepackage{amsfonts}       % blackboard math symbols
\usepackage{amsmath}
\usepackage{amsthm}
\usepackage{bm}
\usepackage{nicefrac}       % compact symbols for 1/2, etc.
\usepackage{microtype}      % microtypography
\usepackage{xspace}      % microtypography
\usepackage{graphicx}
\usepackage{placeins}
\usepackage{hyperref}
\usepackage{cleveref}
\usepackage{caption}
\usepackage{subcaption}
\usepackage{bm}
\usepackage{xcolor}
\usepackage{amsthm}
\usepackage{amsfonts}
\usepackage{stmaryrd}
\usepackage{booktabs} % for professional tables
\usepackage{paralist}
\usepackage{xspace}

\newcommand{\theHalgorithm}{\arabic{algorithm}}

\newcommand{\norm}[1]{\left\lVert#1\right\rVert}
\newcommand{\abs}[1]{\left|#1\right|}
\newcommand{\diag}[1]{\text{diag}\left(#1\right)}
\newcommand{\intint}[1]{\left\llbracket#1\right\rrbracket}
\newtheorem{proposition}{Proposition}[section]
\newcommand{\srm}[1]{\textcolor{red}{{\bf Sam:} #1}}
\newcommand{\tim}[1]{\textcolor{red}{{\bf Tim:} #1}}
\newcommand{\ra}[1]{\textcolor{blue}{{\bf ra:} #1}}
\newcommand{\gl}[1]{\textcolor{violet}{{\bf Gl:} #1}}
\newcommand{\mpv}[1]{\textcolor{green}{{\bf MV:} #1}}
\newcommand{\shrink}{{\bf ShrinkNets}\xspace}
\newcommand{\swl}{{\bf SwitchLayer}\xspace}
\newcommand{\swls}{{\bf SwitchLayers}\xspace}

\title{ShrinkNets: Learning Network Size while Training}

% The \author macro works with any number of authors. There are two
% commands used to separate the names and addresses of multiple
% authors: \And and \AND.
%
% Using \And between authors leaves it to LaTeX to determine where to
% break the lines. Using \AND forces a line break at that point. So,
% if LaTeX puts 3 of 4 authors names on the first line, and the last
% on the second line, try using \AND instead of \And before the third
% author name.

\author{
  Guillaume Leclerc, Manasi Vartak, Raul Castro Fernandez, Tim Kraska, Samuel Madden \\
  Computer Science and Artificial Intelligence Laboratory\\
  Massachusetts Institute of Technology\\
  Cambridge, MA 02145 \\
  \texttt{\{leclerc, mvartak, raulfc, kraska, madden\}@mit.edu} \\
}

\begin{document}
% \nipsfinalcopy is no longer used

\maketitle

\begin{abstract}
As neural networks become widely deployed in different applications and on
different hardware, it has become increasingly important to optimize inference
time and model size along with model accuracy.  Most current techniques
optimize model size, model accuracy and inference time in different stages,
resulting in suboptimal results and computational inefficiency.  In this work,
we propose a new technique called \shrink that optimizes all three of these
metrics at the same time.  Specifically we present a new method to
simultaneously optimize network size and model performance by neuron-level
pruning during training.  Neuron-level pruning not only produces much smaller
networks for a given accuracy threshold but also produces dense weight matrices
that are amenable to efficient inference.  By applying our technique to
convolutional as well as fully connected models, we show that \shrink can
reduce network size by 35X with a 6X improvement in inference time with similar
accuracy as models found by traditional training techniques. 
\end{abstract}

\input{introduction}
\input{related_work}
\input{approach}
\input{impl}
\input{evaluation}
\input{conclusion}
\FloatBarrier
\input{appendix}
\bibliographystyle{unsrt}
\bibliography{custom,general}
\end{document}