### 2018-05-16 paper/NIPS/related_work.tex

%!TEX root=paper.tex
\section{Related Work}

%\begin{itemize}
%  \item post-training compression techniques -- brain damage , 
%  \item group sparsity e.g., \cite{Scardapane2017} and non-parametric neural networks -- 
%  \item training dynamics paper: first overfitting and then randomization?, \gl{Here is the ref, if you can introduce it in the flow \cite{Shwartz-Ziv2017}}
%\end{itemize}

There are several lines of work related to optimizing network structure.

%Given the importance of network structure, many researchers have explored the
%problem of finding the best network structure for a given learning task.  The
%proposed techniques broadly fall into five categories: random search andbrute
%force search, hyperparameter optimization, model compression after training,
%resizing models during training, and automated architecture search methods.

\noindent\textbf{Hyperparameter optimization techniques: }
One way to optimize network architecture is to use 
hyperparameter optimization. Although many methods have been 
proposed for hyperparameter optimization, simple techniques such as randomized
search have been shown to work suprisingly well in practice~\cite{BergstraJAMESBERGSTRA2012,Snoek12}.
%  Brute force search of network sizes is
%also become more practical due to faster and more powerful
%hardware~\cite{molchanov2016pruning}.  
The more complex methods include Bayesian techniques such as~\cite{Snoek12} that
select hyperparameter combinations from uncertain areas of the
hyperparameter space. 
Recently, methods based on bandit algorithms (e.g.
~\cite{li2016hyperband, jamieson2016}) have also been proposed to tune
hyperparameters by quickly discarding 
model configurations that perform badly. 
Although these methods can be used to tune the size of each layer in a network, 
in practice, related work presents limited experimental evidence
regarding this, likely because treating each layer as a hyperparameter
would lead to an excessively large search space.
In contrast, with \shrink, the size of the network can be tuned with 
a single parameter.
% More importantly, none of the hyperparameter optimization methods focuses on
% finding small networks, which is a crucial property of ShrinkNets, necessary to
% achieve good inference times.

%As noted before, all of the above techniques require many tens
%to hundreds of models to be trained, making this process computationally
%inefficient and slow.  More practically, the hyperparameter optimization
%literature does not evaluate their methods on network size and instead focuses
%on optimization hyperparameters such as learning rates and weight decay
%parameters.

\noindent\textbf{Model Compression: }Model compression techniques focus on
reducing the model size \emph{after} training, in contrast to ShrinkNets, which
reduces it \emph{while} training. 
Optimal brain damage~\cite{Cun} identifies connections in a network that are
unimportant and then prunes these connections.
DeepCompression~\cite{han2015deepcompression} takes this one step further and in
addition to pruning connections, it quantizes weights to make inference
extremely efficient.  A different vein of work such as ~\cite{romero2014fitnets,
hinton2015distilling} proposes techniques for distilling a network into a
simpler network or a different model. Because these techniques work after
training, they are orthogonal and complementary to ShrinkNets. Further,
some of these techniques, e.g.,~\cite{Han2015,Cun}, produce sparse matrices that
are not likely to improve inference times even though they reduce network size.
%Unlike our technique which works during
%training, these techiques are used after training and it would be interesting to
%apply them to ShrinkNets as well. 
%\cite{Abadi2016b} share the common goal of
%removing entire blocks of parameter to maintain dense matrices, however their
%method only applies to convolutional layers.

%\noindent\textbf{Auto-ML: } Some work focuses on automatically learning
%model architecture through the use of genetic algorithms and reinforcement
%learning techniques~\cite{DBLP:journals/corr/ZophL16, zoph2017learning}. These
%techniques are focused on learning higher-level architectures (e.g., building
%blocks for neural network architectures). In particular, they require to train
%full models and may take weeks to converge.

%do not
%focus on finding small but well-performing networks for inference, which is the
%goal of ShrinkNets.
%\tim{Argument is not really convincing, but those techniques require to train
%full models and might take weeks to converge. }

\noindent\textbf{Dynamically Sizing Networks: }The techniques closest to our
proposed method are those based on group sparsity such as
~\cite{Scardapane2017}, and those like~\cite{Philipp} that dynamically grow and shrink
 the size of the network during training.  \cite{Scardapane2017}
presents a method that also deactivates neurons using a loss function based on
group-sparsity.  However, the exact details of how their method works are not
given, and their experimental results (on a small, fully connected network), are
substantially worse than ours as shown in Section 5.
\cite{Philipp} propose a method called Adaptive Radial-Angular Gradient Descent
that adds neurons on the fly and removes neurons via an $l_2$ penalty.  However,
this approach requires a new optimizer and takes longer to converge compared to
ShrinkNets.